---
output: pdf_document
header-includes: 
  \usepackage{booktabs}
  \usepackage{microtype}
  \usepackage{longtable}
  \usepackage[margin=1in]{geometry}
---

```{r setup, include=FALSE, results='hide', warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(kableExtra)
library(stringr)
library(dplyr)

linesep<-function(x,y=character()){
  if(!length(x))
    return(y)
  linesep(x[-length(x)], c(rep('',x[length(x)]-1),'\\addlinespace',y))  
}

make_table <- function(tabl, digits = getOption("digits"), BT = TRUE, fontsize = 7, LS){
  # tabl <- readRDS(file) 
  #tabl[tabl == "~"] <- ""
  # if(" " %in% names(tabl)) names(tabl)[names(tabl) == " "] <- "Sig."
  #if(" " %in% names(tabl)) tabl$` ` <- gsub('\\\\', "", tabl$` `)

  if("Q2.5" %in% names(tabl)) tabl <- rename(tabl, `Lower 95% CI` = "Q2.5")
  if("Q97.5" %in% names(tabl)) tabl <- rename(tabl, `Upper 95% CI` = "Q97.5")

  kable(tabl, format = "latex", booktabs = BT, digits = digits, linesep = linesep(LS)) %>%
    kable_styling(font_size = fontsize, full_width = FALSE, latex_options = "hold_position")
}
```


\newpage
\pagenumbering{gobble}

\begin{center}
\huge{\textbf{Online Supplementary Material}}
\end{center}
\vspace{0.5cm}
\textbf{Social immunity and chemical communication in the honeybee: immune-challenged bees enter enforced or self-imposed exile}, by
Tarli Conroy and Luke Holman.
\vspace{0.5cm}



The figures and tables in this document, along with the with the R code used to generate them, can also be viewed online at https://lukeholman.github.io/social_immunity/


\newpage

***Table S1***: Table summarising the posterior estimates of each fixed effect in the Bayesian multinomial logistic model of Experiment 1. Because there were three possible outcomes for each bee (Stayed inside, Left voluntarily, or Forced out), there are two parameter estimates for each predictor in the model. 'Treatment' is a fixed factor with three levels, and the effects shown here are expressed relative to the 'Intact control' group. The $p$ column gives the posterior probability that the true effect size is opposite in sign to what is reported in the Estimate column, similarly to a $p$-value.


```{r, results='asis'}
readRDS("../figures/tableS1.rds") %>%
  mutate(Response = c("% bees leaving voluntarily", "", "", "% bees forced out", "", "")) %>%
  select(Response, everything()) %>%
  make_table(digits = 2, LS = c(3,3)) %>% 
  column_spec(5, width = "1cm") %>%
  column_spec(6, width = "1cm") 
```

\newpage

***Table S2***: This table gives statistics associated with each of the contrasts plotted in Figure 1B. Each pair of rows gives the absolute effect size (i.e. the difference in % bees) and standardised effect size (as log odds ratio; LOR) for the focal treatment, relative to the treatment shown in parentheses, for one of the three possible outcomes (Stayed inside, Left voluntarily, or Forced out). A LOR of $|log(x)|$ indicates that the outcome is $x$ times more frequent in one treatment compared to the other, e.g. $log(2) = 0.69$ and $log(0.5) = -0.69$ correspond to a two-fold difference in frequency. The $PP$ column gives the posterior probability that the true effect size has the same sign as is shown in the Estimate column; this metric has a similar interpretation to a one-tailed $p$ value.

```{r}
readRDS("../figures/tableS2.rds") %>%
  mutate(Response = c("% bees staying inside", rep("", 5), 
                      "% bees leaving voluntarily", rep("", 5),
                      "% bees forced out", rep("", 5))) %>%
  rename(`Est. Error` = Est.Error) %>% 
  select(Response, everything()) %>%
  make_table(digits = 2, LS = c(6,6,6), fontsize = 6) %>% 
  column_spec(5, width = ".8cm") %>%
  column_spec(6, width = ".9cm") %>%
  column_spec(7, width = ".9cm")
```


\newpage

***Table S3***: Table summarising the posterior estimates of each fixed effect in the model of Experiment 2. This was a multinomial model with three possible outcomes (Stayed inside, Left voluntarily, or Forced out), and so there are two parameter estimates for each predictor in the model. 'Treatment' is a fixed factor with two levels, and the effect of LPS shown here is expressed relative to the 'Ringers' treatment. The PP column gives the posterior probability that the true effect size is opposite in sign to what is reported in the Estimate column, similarly to a $p$-value.

```{r, results='asis'}
# make_table("../figures/tableS3.rds", digits = 2) %>%
#   pack_rows("% bees leaving voluntarily", 1, 2) %>%
#   pack_rows("% bees forced out", 3, 4) %>%
#   column_spec(4, width = "1cm") %>%
#   column_spec(5, width = "1cm")

readRDS("../figures/tableS3.rds") %>%
  mutate(Response = c("% bees leaving voluntarily", "", "% bees forced out", "")) %>%
  select(Response, everything()) %>%
  make_table(digits = 2, LS = c(2,2)) %>% 
  column_spec(5, width = "1cm") %>%
  column_spec(6, width = "1cm") 
```
\newpage

***Table S4***: This table gives statistics associated with each of the contrasts plotted in Figure 2B. Each pair of rows gives the absolute (i.e. the difference in % bees) and standardised effect size (as log odds ratio; LOR) for the LPS treatment, relative to the Ringers control, for one of the three possible outcomes (Stayed inside, Left voluntarily, or Forced out). A LOR of $|log(x)|$ indicates that the outcome is $x$ times more frequent in one treatment compared to the other, e.g. $log(2) = 0.69$ indicates a two-fold difference in frequency. The $PP$ column gives the posterior probability that the true effect size has the same sign as is shown in the Estimate column; this metric has a similar interpretation to a one-tailed $p$ value in frequentist statistics.

```{r, results='asis'}
readRDS("../figures/tableS4.rds") %>%
  mutate(Response = c("% bees staying inside", rep("", 1), 
                      "% bees leaving voluntarily", rep("", 1),
                      "% bees forced out", rep("", 1))) %>%
  rename(`Est. Error` = Est.Error) %>% 
  select(Response, everything()) %>%
  make_table(digits = 2, LS = c(2,2,2), fontsize = 6) %>% 
  column_spec(5, width = ".8cm") %>%
  column_spec(6, width = ".9cm") %>%
  column_spec(7, width = ".9cm")
```
\newpage

***Table S5:*** Table summarising the posterior estimates of each fixed effect in the model of Experiment 3 (a binomial GLMM where the response variable was 0 when bees were not in close contact, and 1 when they were). 'Treatment' is a fixed factor with two levels, and the effect of LPS shown here is expressed relative to the 'Ringers' treatment. 'Hive' was a fixed factor with four levels (modelled using deviation coding). The model also included one random effect, 'pair ID', which grouped observations made on each pair of bees, preventing pseudoreplication. The $PP$ column gives the posterior probability that the true effect size is opposite in sign to what is reported in the Estimate column, similarly to a $p$-value.


```{r, results='asis'}
readRDS("../figures/tableS5.rds") %>%
  mutate(` ` = c("***", "*", "", "", "~")) %>% 
  make_table(digits = 2, LS = c(1,1,3))
```

\newpage

***Table S6:*** Pairs in which one bee had received LPS were observed in close contact less frequently than pairs in which one bee had received Ringers. The $PP$ column gives the posterior probability that the true effect size is opposite in sign to what is reported in the Estimate column, similarly to a $p$-value.
```{r, results='asis'}
readRDS("../figures/tableS6.rds") %>%
  mutate(` ` = c("", "*")) %>%
  make_table(digits = 2, LS = c(2))
```

