---
title: "Analysis of Experiment 3"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

## Load data and R packages

```{r results='hide', message=F, warning=F}
# All but 1 of these packages can be easily installed from CRAN.
# However it was harder to install the showtext package. On Mac, I did this:
# installed 'homebrew' using Terminal: ruby -e "$(curl -fsSL https://raw.githubusercontent.com/Homebrew/install/master/install)" 
# installed 'libpng' using Terminal: brew install libpng
# installed 'showtext' in R using: devtools::install_github("yixuan/showtext")  
library(showtext)

library(brms)
library(bayesplot)
library(tidyverse)
library(gridExtra)
library(kableExtra)
library(bayestestR)
library(cowplot)
library(tidybayes)
source("code/helper_functions.R")

# set up nice font for figure
nice_font <- "PT Serif"
font_add_google(name = nice_font, family = nice_font, regular.wt = 400, bold.wt = 700)
showtext_auto()

files <- c("data/data_collection_sheets/hiveA_touching.csv",
           "data/data_collection_sheets/hiveB_touching.csv",
           "data/data_collection_sheets/hiveC_touching.csv",
           "data/data_collection_sheets/hiveD_touching.csv")

experiment3 <- map(
  files, 
  ~ read_csv(.x) %>%
    gather(minute, touching, -tube) %>%
    mutate(treatment = ifelse(substr(tube, 1, 1) %in% c("A", "B"), "AB", "CD"),
           minute = as.numeric(minute),
           touching = as.integer(touching),
           hive = gsub("hive", "", str_extract(.x, "hive[ABCD]")), 
           tube = paste(hive, tube, sep = "_"))
  ) %>% bind_rows() %>%
  mutate(treatment = replace(treatment, 
                             treatment == "AB" & hive %in% c("A", "B"), 
                             "Ringers"),
         treatment = replace(treatment, 
                             treatment == "AB" & hive %in% c("C", "D"), 
                             "LPS"),
         treatment = replace(treatment, 
                             treatment == "CD" & hive %in% c("A", "B"), 
                             "LPS"),
         treatment = replace(treatment, 
                             treatment == "CD" & hive %in% c("C", "D"), 
                             "Ringers")) %>%
  mutate(hive = replace(hive, hive == "A", "Garden"),
         hive = replace(hive, hive == "B", "Skylab"),
         hive = replace(hive, hive == "C", "Arts"),
         hive = replace(hive, hive == "D", "Zoology"))


expt3_counts <- experiment3 %>%
  group_by(treatment, tube, hive) %>%
  summarise(n_touching = sum(touching),
            n_not_touching = sum(touching== 0),
            percent = n_touching / (n_touching + n_not_touching)) %>% 
  ungroup() %>%
  filter(!is.na(n_touching)) %>%
  mutate(hive = C(factor(hive), sum)) # sum coding for hive
```

## Inspect the raw data

### Sample sizes by treatment
```{r}
expt3_counts %>%
  group_by(treatment) %>%
  summarise(n = n()) %>%
  kable() %>% kable_styling(full_width = FALSE)
```

### Sample sizes by treatment and hive
```{r}
expt3_counts %>%
  group_by(hive, treatment) %>%
  summarise(n = n()) %>%
  kable() %>% kable_styling(full_width = FALSE)
```

### Inspect the results 

```{r}
expt3_counts %>%
  group_by(hive, treatment) %>%
  summarise(pc = mean(100 * percent),
            SE = sd(percent) /sqrt(n()),
            n = n()) %>%
  rename(`% observations in which bees were in close contact` = pc,
         Hive = hive, Treatment = treatment) %>% 
  kable(digits = 3) %>% kable_styling(full_width = FALSE) %>%
  column_spec(3, width = "2in")
```

### Plotted as a density histogram

Note that bees more often spent close to 100% of the observation period in contact in the control group, relative to the group treated with LPS.

```{r}
raw_histogram <- expt3_counts %>% 
  filter(treatment == "Ringers") %>%
  ggplot(aes(100 * percent, 
             fill = treatment)) + 
  geom_histogram(data = expt3_counts %>% 
                   filter(treatment == "LPS"), 
                 mapping = aes(y=..density..), 
                 alpha = 0.5, bins = 11, colour = "black", linetype = 2) +
  geom_histogram(mapping = aes(y=..density..), 
                 alpha = 0.5, bins = 11, 
                 colour = "black") + 
  scale_fill_brewer(palette = "Set1", 
                    direction = 1, name = "Treatment") + 
  xlab("% Time in close contact") + ylab("Density") + 
  theme_bw() + 
  theme(legend.position = c(0.27, 0.8),
        text = element_text(family = nice_font)) 

raw_histogram
```



## Binomial model of time spent in contact

### Run the models

Fit a binomial model, where the response is either a 0 (if bees were not in contact) or 1 (if they were). To assess the effects of our predictor variables, we compare 5 models with different fixed factors, ranking them by posterior model probability.

```{r}
# new verison....
if(!file.exists("output/exp3_model.rds")){
  exp3_model_v1 <- brm(
    n_touching | trials(n) ~ treatment * hive + (1 | tube), 
    data = expt3_counts %>% mutate(n = n_touching + n_not_touching), 
    prior = c(set_prior("normal(0, 3)", class = "b")),
    family = "binomial", save_all_pars = TRUE, sample_prior = TRUE,
    chains = 4, cores = 1, iter = 10000, seed = 1)
  
  exp3_model_v2 <- brm(
    n_touching | trials(n) ~ treatment + hive + (1 | tube), 
    data = expt3_counts %>% mutate(n = n_touching + n_not_touching), 
    prior = c(set_prior("normal(0, 3)", class = "b")),
    family = "binomial", save_all_pars = TRUE, sample_prior = TRUE,
    chains = 4, cores = 1, iter = 10000, seed = 1)
  
  exp3_model_v3 <- brm(
    n_touching | trials(n) ~ hive + (1 | tube), 
    data = expt3_counts %>% mutate(n = n_touching + n_not_touching), 
    prior = c(set_prior("normal(0, 3)", class = "b")),
    family = "binomial",  save_all_pars = TRUE, sample_prior = TRUE,
    chains = 4, cores = 1, iter = 10000, seed = 1)
  
  posterior_model_probabilities <- tibble(
    Model = c("treatment * hive + observation_time_minutes",
              "treatment + hive + observation_time_minutes",
              "hive + observation_time_minutes"),
    post_prob = as.numeric(post_prob(exp3_model_v1,
                                     exp3_model_v2,
                                     exp3_model_v3))) %>%
    arrange(-post_prob)
  
  saveRDS(exp3_model_v2, "output/exp3_model.rds") # save the top model, treatment + hive
  saveRDS(posterior_model_probabilities, "output/exp3_post_prob.rds")
}
exp3_model <- readRDS("output/exp3_model.rds")
model_probabilities <- readRDS("output/exp3_post_prob.rds")
```


### Posterior model probabilites

```{r}
model_probabilities %>% 
  kable(digits = 3) %>% kable_styling(full_width = FALSE)
```

### Fixed effects from the top non-null model

```{r}
get_fixed_effects_with_p_values(exp3_model) %>% 
  kable(digits = 3) %>% kable_styling()
```



```{r eval=F}
movement_files <- c("data/data_collection_sheets/hiveA_movement.csv",
                    "data/data_collection_sheets/hiveB_movement.csv",
                    "data/data_collection_sheets/hiveC_movement.csv",
                    "data/data_collection_sheets/hiveD_movement.csv")
movement <- map(1:4, ~ read_csv(movement_files[.x]) %>%
                  mutate(hive = c("A", "B", "C", "D")[.x]) %>%
                  gather(video, movement, -hive, -tube)) %>%
  bind_rows() %>%
  mutate(treatment = ifelse(substr(tube, 1, 1) %in% c("A", "B"), "AB", "CD")) %>%
  mutate(treatment = replace(treatment, treatment == "AB" & hive %in% c("A", "B"), "Ringers"),
         treatment = replace(treatment, treatment == "AB" & hive %in% c("C", "D"), "LPS"),
         treatment = replace(treatment, treatment == "CD" & hive %in% c("A", "B"), "LPS"),
         treatment = replace(treatment, treatment == "CD" & hive %in% c("C", "D"), "Ringers")) %>%
  split(.$hive) %>%
  map(~ .x %>% arrange(video) %>% 
        mutate(video = as.numeric(factor(video, unique(video))))) %>%
  bind_rows() %>%
  mutate(hive = replace(hive, hive == "A", "Garden"),
         hive = replace(hive, hive == "B", "Skylab"),
         hive = replace(hive, hive == "C", "Arts"),
         hive = replace(hive, hive == "D", "Zoology"))

movement %>%
  mutate(movement = factor(movement)) %>%
  group_by(movement, treatment) %>%
  summarise(n=n()) %>%
  ggplot(aes(movement, n, fill = treatment)) +
  geom_bar(stat = "identity", colour = "grey20", position = "dodge") +
  scale_fill_brewer(palette="Pastel2", direction = 1)

library(brms)
mod <- brm(movement | cat(3) ~ treatment + (1 | tube),
           data = movement %>% mutate(movement = movement + 1), family = cumulative())
marginal_effects(mod, categorical = T)

new <- expand.grid(hive = c("Arts", "Garden", "Skylab", "Zoology"),
                   treatment = c("Ringers", "LPS"))

data.frame(new, rbind(fitted(mod, newdata = new, re_formula = NA)[,,1],
                      fitted(mod, newdata = new, re_formula = NA)[,,2],
                      fitted(mod, newdata = new, re_formula = NA)[,,3])) %>%
  mutate(movement = factor(rep(0:2, each = nrow(new)))) %>%
  ggplot(aes(movement, Estimate, ymin = Q2.5, ymax=Q97.5, colour = treatment)) +
  geom_errorbar(position = position_dodge(0.2), width = 0) +
  geom_point(position = position_dodge(0.2)) +
  facet_wrap(~ hive)
```


## Plotting estimates from the model

```{r fig.height=3.2, fig.width=8.6, fig.showtext = TRUE, warning=FALSE}
# new test...
new <- expt3_counts %>% 
  select(treatment) %>% distinct() %>% 
  mutate(n = 100, key = paste("V", 1:n(), sep = ""),
         hive = NA) 
plotting_data <- as.data.frame(fitted(exp3_model, 
                                      newdata=new, re_formula = NA, summary = FALSE))
names(plotting_data) <- c("LPS", "Ringers")
plotting_data <- plotting_data %>% gather(treatment, percent_time_in_contact)

cols <- c("#34558b", "#4ec5a5", "#ffaf12")

dot_plot <- plotting_data %>%
  mutate(treatment = factor(treatment, c("Ringers", "LPS"))) %>%
  ggplot(aes(percent_time_in_contact, treatment)) + 
  stat_dotsh(quantiles = 100, fill = "grey40", colour = "grey40") + 
  stat_pointintervalh(
    mapping = aes(colour = treatment, fill = treatment),
    .width = c(0.5, 0.95), 
    position = position_nudge(y = -0.07),
    point_colour = "grey26", pch = 21, stroke = 0.4) + 
  xlab("Mean % time in close contact") + ylab("Treatment") + 
  scale_colour_brewer(palette = "Pastel1", 
                      direction = -1, name = "Treatment") +
  scale_fill_brewer(palette = "Pastel1", 
                      direction = -1, name = "Treatment") +
  theme_bw() + 
  coord_cartesian(ylim=c(1.4, 2.4)) + 
  theme(
    text = element_text(family = nice_font),
    strip.background = element_rect(fill = "#eff0f1"),
    panel.grid.major.y = element_blank(),
    legend.position = "none"
  ) 
  
# positive effect = odds of this outcome are higher for trt2 than trt1 (put control as trt1)
get_log_odds <- function(trt1, trt2){ 
  log((trt2 / (1 - trt2) / (trt1 / (1 - trt1))))
}

LOR <- plotting_data %>%
  mutate(posterior_sample = rep(1:(n()/2), 2)) %>%
  spread(treatment, percent_time_in_contact) %>%
  mutate(LOR = get_log_odds(Ringers/100, LPS/100)) %>%
  select(LOR)

LOR_plot <- LOR %>%
  ggplot(aes(LOR, y =1)) + 
  geom_vline(xintercept = 0, linetype = 2) +
  stat_dotsh(quantiles = 100, fill = "grey40", colour = "grey40") + 
  stat_pointintervalh(
    colour = "orange", fill = "orange",
    .width = c(0.5, 0.95), 
    position = position_nudge(y = -0.1),
    point_colour = "grey26", pch = 21, stroke = 0.4) + 
  coord_cartesian(ylim=c(0.86, 2)) + 
  xlab("Effect of LPS on % time\nin close contact (LOR)") + 
  ylab("Posterior density") + 
  theme_bw() +
  theme(
    text = element_text(family = nice_font),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    panel.grid.major.y = element_blank(),
    panel.grid.minor.y = element_blank(),
    legend.position = "none"
  ) 
 
p <- cowplot::plot_grid(raw_histogram, 
                        dot_plot, LOR_plot,  labels = c("A", "B", "C"),
                        nrow = 1, align = 'h', axis = 'l')
ggsave(plot = p, filename = "figures/fig3.pdf", height = 3.2, width = 8.6)
p
```
**Figure 3:** Panel A shows the frequency distribution of the % time in close contact, for pairs of bees from the LPS treatment and the Ringers control. Panel B shows the posterior estimates of the mean % time spent in close contact; the details of the quantile dot plot and error bars are the same as described for Figure 1. Panel C shows the effect size (LOR; log odds ratio) associated with the difference in means in Panel B.


## Hypothesis testing and effect sizes

```{r}
get_log_odds <- function(trt1, trt2){ 
  log((trt2 / (1 - trt2) / (trt1 / (1 - trt1))))
}

my_summary <- function(df) {
  
  diff <- (df %>% pull(Ringers)) - (df %>% pull(LPS))
  LOR <- get_log_odds((df %>% pull(Ringers))/100, 
                      (df %>% pull(LPS))/100)
  p <- 1 - (diff %>% bayestestR::p_direction() %>% as.numeric())
  diff <- diff %>% posterior_summary() %>% as_tibble()
  LOR <- LOR %>% posterior_summary() %>% as_tibble()
  output <- rbind(diff, LOR) %>% 
    mutate(p=p, 
           Metric = c("Absolute difference in % time in close contact", 
                      "Log odds ratio")) %>% 
      select(Metric, everything()) %>%
    mutate(p = format(round(p, 4), nsmall = 4))
  output$p[1] <- " "
  output
}

plotting_data %>%
  as_tibble() %>%
  mutate(sample = rep(1:(n() / 2), 2)) %>%
  spread(treatment, percent_time_in_contact) %>%
  mutate(difference = LPS - Ringers) %>%
  my_summary() %>%
  mutate(` ` = ifelse(p < 0.05, "\\*", ""),
         ` ` = replace(` `, p < 0.01, "**"),
         ` ` = replace(` `, p < 0.001, "***"),
         ` ` = replace(` `, p == " ", "")) %>%
  kable(digits = 3) %>% kable_styling() 
```
**Table XX:** Pairs in which one bee had received LPS were observed in close contact less frequently than pairs in which one bee had received Ringers solution. 
