---
title: "Analysis of Experiment 3"
editor_options:
  chunk_output_type: console
output: 
  workflowr::wflow_html:
    code_folding: hide 
---

```{r, include=FALSE}
options(width=120)
local({
  hook_output <- knitr::knit_hooks$get('output')
  knitr::knit_hooks$set(output = function(x, options) {
    options$attr.output <- c(
      options$attr.output,
      sprintf('style="max-height: %s;"', options$max.height)
    )
    hook_output(x, options)
  })
})
```

## Load data and R packages

```{r results='hide', message=F, warning=F}
# All but 1 of these packages can be easily installed from CRAN.
# However it was harder to install the showtext package. On Mac, I did this:
# installed 'homebrew' using Terminal: ruby -e "$(curl -fsSL https://raw.githubusercontent.com/Homebrew/install/master/install)" 
# installed 'libpng' using Terminal: brew install libpng
# installed 'showtext' in R using: devtools::install_github("yixuan/showtext")  
library(showtext)

library(brms)
library(bayesplot)
library(tidyverse)
library(gridExtra)
library(kableExtra)
library(bayestestR)
library(cowplot)
library(tidybayes)
library(scales)
library(car)
source("code/helper_functions.R")

# set up nice font for figure
nice_font <- "Lora"
font_add_google(name = nice_font, family = nice_font, regular.wt = 400, bold.wt = 700)
showtext_auto()

experiment3 <- read_csv("data/clean_data/experiment_3.csv")

expt3_counts <- experiment3 %>%
  group_by(treatment, pairID, hive) %>%
  summarise(n_touching = sum(touching),
            n_not_touching = sum(touching== 0),
            percent = n_touching / (n_touching + n_not_touching), 
            .groups = "drop") %>% 
  ungroup() %>%
  filter(!is.na(n_touching)) %>%
  mutate(treatment = factor(treatment, c("Ringers", "LPS"))) %>%
  mutate(hive = C(factor(hive), sum)) # sum coding for hive
```


## Inspect the raw data {.tabset .tabset-fade}

### Sample sizes by treatment
```{r}
sample_sizes <- expt3_counts %>%
  group_by(treatment) %>%
  summarise(n = n(), .groups = "drop") 

sample_sizes %>%
  kable() %>% kable_styling(full_width = FALSE)
```

### Sample sizes by treatment and hive
```{r}
expt3_counts %>%
  group_by(hive, treatment) %>%
  summarise(n = n(), .groups = "drop") %>%
  kable() %>% kable_styling(full_width = FALSE)
```

### Means and standard errors

```{r}
expt3_counts %>%
  group_by(hive, treatment) %>%
  summarise(pc = mean(100 * percent),
            SE = sd(percent) /sqrt(n()),
            n = n(), .groups = "drop") %>%
  rename(`% observations in which bees were in close contact` = pc,
         Hive = hive, Treatment = treatment) %>% 
  kable(digits = 3) %>% kable_styling(full_width = FALSE) %>%
  column_spec(3, width = "2in")
```

## Histogram of the results

Note that bees more often spent close to 100% of the observation period in contact in the control group, relative to the group treated with LPS.

```{r fig.showtext = TRUE}
histo_data <- expt3_counts %>%
  left_join(sample_sizes, by = "treatment") %>%
  arrange(treatment)

raw_histogram <- histo_data %>% 
  filter(grepl("Ringers", treatment)) %>%
  ggplot(aes(100 * percent, 
             fill = treatment)) + 
  geom_histogram(data = histo_data %>% 
                   filter(grepl("LPS", treatment)), 
                 mapping = aes(y = ..density..), 
                 alpha = 0.5, bins = 11, colour = "black", linetype = 2) +
  geom_histogram(mapping = aes(y = ..density..), 
                 alpha = 0.5, bins = 11, 
                 colour = "black") + 
  scale_fill_brewer(palette = "Set1", 
                    direction = 1, name = "Treatment") + 
  xlab("% Time in close contact") + ylab("Density") + 
  theme_bw() + 
  theme(legend.position = c(0.17, 0.82),
        legend.background = element_rect(fill = scales::alpha('white', 0.7)),
        text = element_text(family = nice_font)) 

rm(histo_data)
raw_histogram
```




## Bayesian binomial model of time spent in contact

### Run the models

Fit three different binomial models, where the response is either a 0 (if bees were not in contact) or 1 (if they were). 

```{r}
if(!file.exists("output/exp3_model.rds")){
  
  prior <- c(set_prior("normal(0, 3)", class = "b"),
             set_prior("normal(0, 1)", class = "sd", group = "pairID"))
  
  exp3_model <- brm(
    n_touching | trials(n) ~ treatment + hive + (1 | pairID), 
    data = expt3_counts %>% mutate(n = n_touching + n_not_touching), 
    prior = prior,
    family = "binomial", 
    chains = 4, cores = 1, iter = 20000, seed = 1)
  
  saveRDS(exp3_model, "output/exp3_model.rds") 
}
exp3_model <- readRDS("output/exp3_model.rds")
```


### Posterior predictive check

This plot shows ten predictions from the posterior (pale blue) as well as the original data (dark blue), for the proportion of time spent in close contact (0-100%). The predicted time in contact is similar to the real data, illustrating that the model is able to recapitulate the original data fairly closely (a necessary requirement for making inferences from the model).

```{r message=F, warning=F}
pp_check(exp3_model, nsamples = 20)
```

### Fixed effects from the top model

#### Raw output of the `treatment + hive` model

```{r max.height='300px', max.width='400px'}
summary(exp3_model)
```

#### Formatted `brms` output for Table S5
The code chunk below wrangles the raw output of the `summary()` function for `brms` models into a more readable table of results, and also adds 'Bayesian p-values' (i.e. the posterior probability that the true effect size has the same sign as the reported effect). 

******

***Table S5:*** Table summarising the posterior estimates of each fixed effect in the best-fitting model of Experiment 3 that contained the treatment effect. This was a binomial model where the response variable was 0 for observations in which bees were not in close contact, and 1 when they were. 'Treatment' is a fixed factor with two levels, and the effect of LPS shown here is expressed relative to the 'Ringers' treatment. 'Hive' was a fixed factor with four levels; unlike for treatment, we modelled hive using deviation coding, such that the intercept term represents the mean across all hives (in the Ringers treatment), and the three hive terms represent the deviation from this mean for three of the four hives. The model also included one random effect, 'pair ID', which grouped observations made on each pair of bees, preventing pseudoreplication. The $PP$ column gives the posterior probability that the true effect size is opposite in sign to what is reported in the Estimate column, similarly to a $p$-value.

```{r}
tableS5 <- get_fixed_effects_with_p_values(exp3_model) %>% 
  mutate(mu = map_chr(str_extract_all(Parameter, "mu[:digit:]"), ~ .x[1]),
         Parameter = str_remove_all(Parameter, "mu[:digit:]_"),
         Parameter = str_replace_all(Parameter, "treatment", "Treatment: "),
         Parameter = str_replace_all(Parameter, "observation_time_minutes", "Observation duration (minutes)")) %>%
  arrange(mu) %>%
  select(-mu, -Rhat, -Bulk_ESS, -Tail_ESS) %>%
    mutate(PP = format(round(PP, 4), nsmall = 4))

names(tableS5)[3:5] <- c("Est. Error", "Lower 95% CI", "Upper 95% CI")

saveRDS(tableS5, file = "figures/tableS5.rds")

tableS5 %>%
  kable(digits = 3) %>% 
  kable_styling(full_width = FALSE) 
```




## Plotting estimates from the model

```{r fig.height=3.2, fig.width=8.6, fig.showtext = TRUE, warning=FALSE}
new <- expt3_counts %>% 
  select(treatment) %>% distinct() %>% 
  mutate(n = 100, key = paste("V", 1:n(), sep = ""),
         hive = NA) 
plotting_data <- as.data.frame(
  fitted(exp3_model, 
         newdata=new, re_formula = NA, summary = FALSE))
names(plotting_data) <- c("LPS", "Ringers")
plotting_data <- plotting_data %>% gather(treatment, percent_time_in_contact)

panel_c_colour <- "#CC79A7"

dot_plot <- plotting_data %>%
  mutate(treatment = factor(treatment, c("Ringers", "LPS"))) %>%
  ggplot(aes(percent_time_in_contact, treatment)) + 
  stat_dotsh(quantiles = 100, fill = "grey40", colour = "grey40") + 
  stat_pointintervalh(
    mapping = aes(colour = treatment, fill = treatment),
    .width = c(0.5, 0.95), 
    position = position_nudge(y = -0.07),
    point_colour = "grey26", pch = 21, stroke = 0.4) + 
  xlab("Mean % time in close contact") + ylab("Treatment") + 
  scale_colour_brewer(palette = "Pastel1", 
                      direction = -1, name = "Treatment") +
  scale_fill_brewer(palette = "Pastel1", 
                      direction = -1, name = "Treatment") +
  theme_bw() + 
  coord_cartesian(ylim=c(1.4, 2.4)) + 
  theme(
    text = element_text(family = nice_font),
    strip.background = element_rect(fill = "#eff0f1"),
    panel.grid.major.y = element_blank(),
    legend.position = "none"
  ) 
  
# positive effect = odds of this outcome are higher for trt2 than trt1 (put control as trt1)
get_log_odds <- function(trt1, trt2){ 
  log((trt2 / (1 - trt2) / (trt1 / (1 - trt1))))
}

# LOR <- plotting_data %>%
#   mutate(posterior_sample = rep(1:(n()/2), 2)) %>%
#   spread(treatment, percent_time_in_contact) %>%
#   mutate(LOR = get_log_odds(Ringers/100, LPS/100)) %>%
#   select(LOR)
# 
# LOR_plot <- LOR %>%
#   ggplot(aes(LOR, y =1)) + 
#   geom_vline(xintercept = 0, linetype = 2) +
#   stat_dotsh(quantiles = 100, fill = "grey40", colour = "grey40") + 
#   stat_pointintervalh(
#     colour = panel_c_colour, fill = panel_c_colour,
#     .width = c(0.5, 0.95), 
#     position = position_nudge(y = -0.1),
#     point_colour = "grey26", pch = 21, stroke = 0.4) + 
#   coord_cartesian(ylim=c(0.86, 2)) + 
#   xlab("Effect of LPS on mean\n% time in close contact (LOR)") + 
#   ylab("Posterior density") + 
#   theme_bw() +
#   theme(
#     text = element_text(family = nice_font),
#     axis.text.y = element_blank(),
#     axis.ticks.y = element_blank(),
#     panel.grid.major.y = element_blank(),
#     panel.grid.minor.y = element_blank(),
#     legend.position = "none"
#   ) 
 
p <- cowplot::plot_grid(raw_histogram, 
                        dot_plot, labels = c("A", "B"),
                        nrow = 1, align = 'h', axis = 'l')
ggsave(plot = p, filename = "figures/fig3.pdf", height = 3.4, width = 9)
p
```
***Figure 3:*** Results of Experiment 3 (n = 439 bees). Panel A shows the frequency distribution of the % time in close contact, for pairs of bees from the LPS treatment and the Ringers control. Panel B shows the posterior estimates of the mean % time spent in close contact; the details of the quantile dot plot and error bars are the same as described for Figure 1. 




## Hypothesis testing and effect sizes

***Table S6:*** Pairs in which one bee had received LPS were observed in close contact less frequently than pairs in which one bee had received Ringers. 

```{r}
get_log_odds <- function(trt1, trt2){ 
  log((trt2 / (1 - trt2) / (trt1 / (1 - trt1))))
}

my_summary <- function(df) {
  
  diff <- (df %>% pull(Ringers)) - (df %>% pull(LPS))
  LOR <- get_log_odds((df %>% pull(Ringers))/100, 
                      (df %>% pull(LPS))/100)
  p <- 1 - (diff %>% bayestestR::p_direction() %>% as.numeric())
  diff <- diff %>% posterior_summary() %>% as_tibble()
  LOR <- LOR %>% posterior_summary() %>% as_tibble()
  output <- rbind(diff, LOR) %>% 
    mutate(PP=p, 
           Metric = c("Absolute difference in % time in close contact", 
                      "Log odds ratio")) %>% 
      select(Metric, everything()) %>%
    mutate(PP = format(round(PP, 4), nsmall = 4))
  output$PP[1] <- " "
  output
}

tableS6 <- plotting_data %>%
  as_tibble() %>%
  mutate(sample = rep(1:(n() / 2), 2)) %>%
  spread(treatment, percent_time_in_contact) %>%
  mutate(difference = LPS - Ringers) %>%
  my_summary() %>%
  mutate(` ` = ifelse(PP < 0.05, "\\*", ""),
         ` ` = replace(` `, PP < 0.01, "**"),
         ` ` = replace(` `, PP < 0.001, "***"),
         ` ` = replace(` `, PP == " ", "")) 

saveRDS(tableS6, "figures/tableS6.rds")

tableS6 %>%
  kable(digits = 3) %>% kable_styling() 
```

