---
title: "Analysis of Experiment 2"
editor_options:
  chunk_output_type: console
output: 
  workflowr::wflow_html:
    code_folding: hide 
---

<!-- ```{r setup, include=FALSE} -->
<!-- library(thematic) -->
<!-- bootstraplib::bs_theme_new() -->
<!-- bootstraplib::bs_theme_fonts("Pacifico") -->

<!-- thematic_on( -->
<!--   bg = "black", fg = "red", accent = "orange" -->
<!-- ) -->
<!-- knitr::opts_chunk$set(echo = FALSE) -->
<!-- ``` -->


```{r, include=FALSE}
options(width=120)
local({
  hook_output <- knitr::knit_hooks$get('output')
  knitr::knit_hooks$set(output = function(x, options) {
    options$attr.output <- c(
      options$attr.output,
      sprintf('style="max-height: %s;"', options$max.height)
    )
    hook_output(x, options)
  })
})
```

## Load data and R packages

```{r results='hide', message=F, warning=F}
# All but 1 of these packages can be easily installed from CRAN.
# However it was harder to install the showtext package. On Mac, I did this:
# installed 'homebrew' using Terminal: ruby -e "$(curl -fsSL https://raw.githubusercontent.com/Homebrew/install/master/install)" 
# installed 'libpng' using Terminal: brew install libpng
# installed 'showtext' in R using: devtools::install_github("yixuan/showtext")  
library(showtext)

library(brms)
library(bayesplot)
library(tidyverse)
library(gridExtra)
library(kableExtra)
library(bayestestR)
library(tidybayes)
library(cowplot)
library(car)
source("code/helper_functions.R")

# set up nice font for figure
nice_font <- "Lora"
font_add_google(name = nice_font, family = nice_font, regular.wt = 400, bold.wt = 700)
showtext_auto()

exp2_treatments <- c("Ringers", "LPS")

durations <- read_csv("data/data_collection_sheets/experiment_durations.csv") %>%
  filter(experiment == 2) %>% select(-experiment)

outcome_tally <- read_csv(file = "data/clean_data/experiment_2_outcome_tally.csv") %>%
  mutate(
    outcome = str_replace_all(outcome, "Stayed inside the hive", "Stayed inside"),
    outcome = str_replace_all(outcome, "Left of own volition", "Left voluntarily"),
    outcome = factor(outcome, levels = c("Stayed inside", "Left voluntarily", "Forced out")),
         treatment = str_replace_all(treatment, "Ringer CHC", "Ringers"),
         treatment = str_replace_all(treatment, "LPS CHC", "LPS"),
         treatment = factor(treatment, levels = exp2_treatments))

# Re-formatted version of the same data, where each row is an individual bee. We need this format to run the brms model.
data_for_categorical_model <- outcome_tally %>%
  mutate(id = 1:n()) %>%
  split(.$id) %>%
  map(function(x){
    if(x$n[1] == 0) return(NULL)
    data.frame(
      treatment = x$treatment[1],
      hive = x$hive[1],
      colour = x$colour[1],
      outcome = rep(x$outcome[1], x$n))
  }) %>% do.call("rbind", .) %>% as_tibble() %>%
  arrange(hive, treatment) %>%
  mutate(outcome_numeric = as.numeric(outcome),
         hive = as.character(hive),
         treatment = factor(treatment, levels = exp2_treatments)) %>%
  left_join(durations, by = "hive") %>%
  mutate(hive = C(factor(hive), sum))  # use "sum coding" for hive, since there is no obvious reference level
```


## Inspect the raw data {.tabset .tabset-fade}

Click the three tabs to see each table.

### Sample sizes by treatment
```{r}
sample_sizes <- data_for_categorical_model %>%
  group_by(treatment) %>%
  summarise(n = n(), .groups = "drop") 

sample_sizes %>%
  kable() %>% kable_styling(full_width = FALSE)
```


### Sample sizes by treatment and hive
```{r}
data_for_categorical_model %>%
  group_by(hive, treatment) %>%
  summarise(n = n(), .groups = "drop") %>%
  spread(treatment, n) %>%
  kable() %>% kable_styling(full_width = FALSE)
```


### Oberved outcomes

```{r}
outcome_tally %>%
  select(-colour) %>% 
  spread(outcome, n) %>%
  kable(digits = 3) %>% kable_styling(full_width = FALSE) 
```

## Plot showing raw means and 95% CI of the mean

```{r fig.width=9, fig.height=5, fig.showtext=TRUE}
pd <- position_dodge(.3)
outcome_tally %>%
  group_by(treatment, outcome) %>%
  summarise(n = sum(n), .groups = "drop") %>% mutate() %>%
  group_by(treatment) %>%
  mutate(total_n = sum(n),
         percent = 100 * n / sum(n),
         SE = sqrt(total_n * (percent/100) * (1-(percent/100)))) %>% 
  ungroup() %>%
  mutate(lowerCI = map_dbl(1:n(), ~ 100 * binom.test(n[.x], total_n[.x])$conf.int[1]),
         upperCI = map_dbl(1:n(), ~ 100 * binom.test(n[.x], total_n[.x])$conf.int[2])) %>%
  filter(outcome != "Stayed inside") %>%
  ggplot(aes(treatment, percent, fill = outcome)) + 
  geom_errorbar(aes(ymin=lowerCI, ymax=upperCI), position = pd, width = 0) + 
  geom_point(stat = "identity", position = pd, colour = "grey15", pch = 21, size = 4) + 
  scale_fill_brewer(palette = "Pastel1", name = "Outcome", direction = -1) + 
  xlab("Treatment") + ylab("% bees (\u00B1 95% CIs)") + 
  theme_bw(20) + 
  theme(text = element_text(family = nice_font),
        legend.position = "top") + 
  coord_flip()
```



## Run the preliminary GLM

The multinomial model below is not commonly used, though we believe it is the right choice for this particular experiment (e.g. because it can model a three-item categorical response variable, and it can incorporate priors). However during peer review, we were asked whether the results were similar when using standard statistical methods. To address this question, we here present a frequentist Generalised Linear Model (GLM; using `base::glm`), which tests the null hypothesis that the proportion of bees exiting the hive (i.e. the proportion leaving voluntarily plus those that were forced out) is equal between treatment groups and hives. The model uses the quasibinomial family, since this family is more robust to overdispersion than the binomial family (though testing revealed both families gave qualitatively identical results).

The model results are similar to those from the multinomial model: bees treated with the CHCs from LPS-treated nestmates left the hive more often than controls, and there were differences between hives in the proportion of bees leaving.

### Run the GLM
```{r message=FALSE, warning=FALSE}
test_dat <- outcome_tally %>% 
  group_by(hive, treatment, outcome) %>% 
  summarise(n = sum(n)) %>% 
  mutate(left = ifelse(outcome == "Stayed inside", "stayed_inside" ,"left_hive")) %>% 
  group_by(hive, treatment, left) %>% 
  summarise(n = sum(n)) %>%
  spread(left, n)  

simple_model <- glm(
    cbind(left_hive, stayed_inside) ~ treatment + hive, 
    data = test_dat, 
    family = "quasibinomial")
```


### Inspect model coefficients and p-values

The reference level for `treatment` is the the Ringers Control, and the reference for `hive` is the hive named Arts.
```{r}
summary(simple_model)
```

### Inspect Type II Anova table
```{r}
Anova(simple_model, test = "LR", Type = "II")
```



## Baysian multinomial model

### Fit the model

Fit the multinomial logisitic models, with a 3-item response variable describing what happened to each bee introduced to the hive: stayed inside, left voluntarily, or forced out by the other workers. 

```{r}
if(!file.exists("output/exp2_model.rds")){

  exp2_model <- brm(
    outcome_numeric ~ treatment + hive + observation_time_minutes, 
    data = data_for_categorical_model, 
    prior = c(set_prior("normal(0, 3)", class = "b", dpar = "mu2"),
              set_prior("normal(0, 3)", class = "b", dpar = "mu3")),
    family = "categorical", save_all_pars = TRUE, sample_prior = TRUE,
    chains = 4, cores = 1, iter = 5000, seed = 1)
  
  saveRDS(exp2_model, "output/exp2_model.rds")
}

exp2_model <- readRDS("output/exp2_model.rds")
```


### Parameter estimates from the model

#### Raw output of the `treatment + hive` model

```{r max.height='300px', max.width='400px'}
summary(exp2_model)
```

#### Formatted `brms` output for Table S3
The code chunk below wrangles the raw output of the `summary()` function for `brms` models into a more readable table of results, and also adds 'Bayesian p-values' (i.e. the posterior probability that the true effect size has the same sign as the reported effect). 

******

***Table S3:*** Table summarising the posterior estimates of each fixed effect in the best-fitting model of Experiment 2. This was a multinomial model with three possible outcomes (stay inside, leave voluntarily, be forced out), and so there are two parameter estimates for the intercept and for each predictor in the model. 'Treatment' is a fixed factor with two levels, and the effect of LPS shown here is expressed relative to the 'Ringers' treatment. 'Hive' was a fixed factor with four levels; unlike for treatment, we modelled hive using deviation coding, such that the intercept term represents the mean across all hives (in the Ringers treatment), and the three hive terms represent the deviation from this mean for three of the four hives. Lastly, observation duration was a continuous predictor expressed to the nearest minute. The PP column gives the posterior probability that the true effect size is opposite in sign to what is reported in the Estimate column, similarly to a $p$-value.

```{r}
tableS3 <- get_fixed_effects_with_p_values(exp2_model) %>% 
  mutate(mu = map_chr(str_extract_all(Parameter, "mu[:digit:]"), ~ .x[1]),
         Parameter = str_remove_all(Parameter, "mu[:digit:]_"),
         Parameter = str_replace_all(Parameter, "treatment", "Treatment: "),
         Parameter = str_replace_all(Parameter, "observation_time_minutes", "Observation duration (minutes)")) %>%
  arrange(mu) %>%
  select(-mu, -Rhat, -Bulk_ESS, -Tail_ESS) %>%
  mutate(PP = format(round(PP, 4), nsmall = 4))

names(tableS3)[3:5] <- c("Est. Error", "Lower 95% CI", "Upper 95% CI")

saveRDS(tableS3, file = "figures/tableS3.rds")

tableS3 %>%
  kable(digits = 3) %>% 
  kable_styling(full_width = FALSE) %>%
  pack_rows("% bees leaving voluntarily", 1, 6) %>%
  pack_rows("% bees forced out", 7, 12)
```



## Plotting estimates from the model

### Derive prediction from the posterior

```{r}
get_posterior_preds <- function(focal_hive){
  new <- expand.grid(treatment = levels(data_for_categorical_model$treatment), 
                     hive = focal_hive,
                     observation_time_minutes = 120) 
  
  preds <- fitted(exp2_model, newdata = new, summary = FALSE)
  dimnames(preds) <- list(NULL, paste(new$treatment, new$hive, sep = "~"), NULL)
  
  rbind(
    as.data.frame(preds[,, 1]) %>% 
      mutate(outcome = "Stayed inside", posterior_sample = 1:n()),
    as.data.frame(preds[,, 2]) %>% 
      mutate(outcome = "Left voluntarily", posterior_sample = 1:n()),
    as.data.frame(preds[,, 3]) %>% 
      mutate(outcome = "Forced out", posterior_sample = 1:n())) %>%
    gather(treatment, prop, contains("~")) %>%
    mutate(treatment = strsplit(treatment, split = "~"),
           hive = map_chr(treatment, ~ .x[2]),
           treatment = map_chr(treatment, ~ .x[1]),
           treatment = factor(treatment, c("Ringers", "LPS")),
           outcome = factor(outcome, c("Stayed inside", "Left voluntarily", "Forced out"))) %>%
    arrange(treatment, outcome) %>% as_tibble() %>% select(-hive)
}


# plotting data for panel A: one specific hive
plotting_data <- get_posterior_preds(focal_hive = "Zoology")

# stats data: for comparing means across all hives
stats_data <- get_posterior_preds(focal_hive = NA)
```

### Make Figure 2

```{r fig.height=6, fig.width=6, fig.showtext = TRUE, warning=FALSE}
cols <- RColorBrewer::brewer.pal(3, "Set2")
panel_c_colour <- "#CC79A7"

dot_plot <- plotting_data %>%
  left_join(sample_sizes, by = "treatment") %>%
  arrange(treatment) %>%
  mutate(treatment = factor(paste(treatment, "\n(n = ", n, ")", sep = ""),
                            unique(paste(treatment, "\n(n = ", n, ")", sep = "")))) %>% 
  ggplot(aes(100 * prop, treatment)) + 
  stat_dotsh(quantiles = 100, fill = "grey40", colour = "grey40") + 
  stat_pointintervalh(aes(colour = outcome, fill = outcome), 
                      .width = c(0.5, 0.95),
                      position = position_nudge(y = -0.07), point_colour = "grey26", pch = 21, stroke = 0.4) + 
  scale_colour_manual(values = cols) + 
  scale_fill_manual(values = cols) + 
  facet_wrap( ~ outcome, scales = "free_x") + 
  xlab("% bees (posterior estimate)") + ylab("Treatment") + 
  theme_bw() + 
  coord_cartesian(ylim=c(1.4, 2.2)) + 
  theme(
    text = element_text(family = nice_font),
    strip.background = element_rect(fill = "#eff0f1"),
    panel.grid.major.y = element_blank(),
    legend.position = "none"
  ) 
  

get_log_odds <- function(trt1, trt2){ # positive effect = odds of this outcome are higher for trt2 than trt1 (put control as trt1)
  log((trt2 / (1 - trt2) / (trt1 / (1 - trt1))))
}

LOR <- plotting_data %>%
  spread(treatment, prop) %>%
  mutate(LOR = get_log_odds(Ringers, LPS)) %>%
  select(posterior_sample, outcome, LOR)



LOR_plot <- LOR %>%
  ggplot(aes(LOR, outcome, colour = outcome)) + 
  geom_vline(xintercept = 0, size = 0.3, colour = "grey20") + 
  geom_vline(xintercept = log(2), linetype = 2, size = 0.6, colour = "grey") +
  geom_vline(xintercept = -log(2), linetype = 2, size = 0.6, colour = "grey") +
  stat_pointintervalh(aes(colour = outcome, fill = outcome), 
                      position = position_dodge(0.4), 
                      .width = c(0.5, 0.95),
                      point_colour = "grey26", pch = 21, stroke = 0.4) + 
  scale_colour_manual(values = cols) + 
  scale_fill_manual(values = cols) + 
  xlab("Effect size of LPS\n(log odds ratio)") + ylab("Mode of exit") + 
  theme_bw() +
  theme(
    text = element_text(family = nice_font),
    panel.grid.major.y = element_blank(),
    legend.position = "none"
  ) 


diff_in_forced_out_plot <- plotting_data %>%
  spread(outcome, prop) %>%
  mutate(prop_leavers_that_were_forced_out = `Forced out` / (`Forced out` + `Left voluntarily`)) %>%
  select(posterior_sample, treatment, prop_leavers_that_were_forced_out)  %>%
  spread(treatment, prop_leavers_that_were_forced_out) %>% 
  mutate(difference_prop_forced_out_LOR = get_log_odds(Ringers, LPS)) %>%
  ggplot(aes(difference_prop_forced_out_LOR, y =1)) + 
  geom_vline(xintercept = 0, linetype = 2) +
  stat_dotsh(quantiles = 100, fill = "grey40", colour = "grey40") + 
  stat_pointintervalh(
    colour = panel_c_colour, fill = panel_c_colour,
    .width = c(0.5, 0.95), 
    position = position_nudge(y = -0.1),
    point_colour = "grey26", pch = 21, stroke = 0.4) + 
  coord_cartesian(ylim=c(0.86, 2)) + 
  xlab("Effect of LPS on proportion\nbees leaving by force\n(log odds ratio)") + 
  ylab("Posterior density") + 
  theme_bw() +
  theme(
    text = element_text(family = nice_font),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    panel.grid.major.y = element_blank(),
    panel.grid.minor.y = element_blank(),
    legend.position = "none"
  ) 
 
# 
# diff_in_forced_out_plot <- plotting_data %>%
#   spread(outcome, prop) %>%
#   mutate(prop_leavers_that_were_forced_out = `Forced out` / (`Forced out` + `Left voluntarily`)) %>%
#   select(posterior_sample, treatment, prop_leavers_that_were_forced_out)  %>%
#   spread(treatment, prop_leavers_that_were_forced_out) %>% 
#   mutate(difference_prop_forced_out_LOR = get_log_odds(Ringers, LPS)) %>%
#   ggplot(aes(difference_prop_forced_out_LOR, y =1)) + 
#   geom_vline(xintercept = 0, linetype = 2) +
#   stat_dotsh(quantiles = 100, fill = "grey40", colour = "grey40") + 
#   stat_pointintervalh(
#     colour = "#0CE3AC", fill = "#0CE3AC",
#     .width = c(0.5, 0.95), 
#     position = position_nudge(y = -0.1),
#     point_colour = "grey26", pch = 21, stroke = 0.4) +
#   xlab("Effect of LPS on proportion\nbees leaving by force\n(log odds ratio)") + 
#   ylab("Posterior density") + 
#   theme_bw() +
#   theme(
#     text = element_text(family = nice_font),
#     panel.grid.major.y = element_blank(),
#     legend.position = "none"
#   ) 

bottom_row <- cowplot::plot_grid(
  LOR_plot, diff_in_forced_out_plot, 
  labels = c("B", "C"),
  nrow = 1, align = 'hv', 
  axis = 'l'
)

top_row <- cowplot::plot_grid(dot_plot, labels = "A")
p <- cowplot::plot_grid(top_row, bottom_row, 
                        nrow = 2, align = 'v', axis = 'l', 
                        rel_heights = c(1.4, 1))
ggsave(plot = p, filename = "figures/fig2.pdf", height = 6, width = 6)
p
```

***Figure 2:*** Panel A shows the same information as Figure 1A. Panel B gives the posterior estimates of the effect size (log odds ratio) of the LPS treatment as a log odds ratio, for each of the three possible outcomes; the details are the same as in Figure 1B. Panel C shows the posterior estimate of the effect of the LPS treatment on the proportion of bees observed leaving the hive by force, as opposed to leaving voluntarily.



## Hypothesis testing and effect sizes

### Posterior effect size estimates

This section calculates the effect size and 95% CIs that are shown in Figure 2B (and creates Table S4). 


***Table S4***: This table gives statistics associated with each of the contrasts plotted in Figure 2B. Each pair of rows gives the absolute (i.e. the difference in % bees) and standardised effect size (as log odds ratio; LOR) for the LPS treatment, relative to the Ringers control, for one of the three possible outcomes (stayed inside, left voluntarily, or forced out). A LOR of $|log(x)|$ indicates that the outcome is $x$ times more frequent in one treatment compared to the other, e.g. $log(2) = 0.69$ indicates a two-fold difference in frequency. The $PP$ column gives the posterior probability that the true effect size has the same sign as is shown in the Estimate column; this metric has a similar interpretation to a one-tailed $p$ value in frequentist statistics. 

```{r}
my_summary <- function(df, columns, outcome) {
  lapply(columns, function(x){
    
    p <- 1 - (df %>% pull(!! x) %>%
                bayestestR::p_direction() %>% as.numeric())
    
    df %>% pull(!! x) %>% posterior_summary() %>% as_tibble() %>% 
      mutate(PP = p, Outcome = outcome, Metric = x) %>% 
      select(Outcome, Metric, everything())
  }) %>% do.call("rbind", .)
}

stats_table <- rbind(
  plotting_data %>%
    filter(outcome == "Stayed inside") %>%
    spread(treatment, prop) %>%
    mutate(`Absolute difference in % bees staying inside` = 100 * (LPS - Ringers),
           `Log odds ratio` = get_log_odds(Ringers, LPS)) %>%
    my_summary(c("Absolute difference in % bees staying inside", 
                 "Log odds ratio"),
               outcome = "Stayed inside") %>%
    mutate(PP = c(" ", format(round(PP[2], 4), nsmall = 4))),
  
  plotting_data %>%
    filter(outcome == "Left voluntarily") %>%
    spread(treatment, prop) %>%
    mutate(`Absolute difference in % bees leaving voluntarily` = 100 * (LPS - Ringers),
           `Log odds ratio` = get_log_odds(Ringers, LPS)) %>%
    my_summary(c("Absolute difference in % bees leaving voluntarily", 
                 "Log odds ratio"),
               outcome = "Left voluntarily") %>%
    mutate(PP = c(" ", format(round(PP[2], 4), nsmall = 4))),
  
  plotting_data %>%
    filter(outcome == "Forced out") %>%
    spread(treatment, prop) %>%
    mutate(`Absolute difference in % bees forced out` = 100 * (LPS - Ringers),
           `Log odds ratio` = get_log_odds(Ringers, LPS)) %>%
    my_summary(c("Absolute difference in % bees forced out", 
                 "Log odds ratio"),
               outcome = "Forced out") %>%
    mutate(PP = c(" ", format(round(PP[2], 4), nsmall = 4))) 
) %>%
  mutate(` ` = ifelse(PP < 0.05, "\\*", ""),
         ` ` = replace(` `, PP < 0.01, "**"),
         ` ` = replace(` `, PP < 0.001, "***"),
         ` ` = replace(` `, PP == " ", ""))

stats_table[c(2,4,6), 1] <- " "

stats_table %>%
  select(-Outcome) %>% saveRDS("figures/tableS4.rds")


stats_table %>%
  select(-Outcome) %>%
  kable(digits = 3) %>% kable_styling(full_width = FALSE) %>% 
  row_spec(c(0,2,4,6), extra_css = "border-bottom: solid;") %>%
  pack_rows("% bees staying inside", 1, 2) %>%
  pack_rows("% bees leaving voluntarily", 3, 4) %>%
  pack_rows("% bees forced out", 5, 6)
```



### Evaluating evidence for the null hypothesis

Here, we derive the result present in prose in the Results, where we calculated the posterior probability that the true effect size lies in the range $-log(2) < LOR < log(2)$.

```{r}
LOR_forced <- LOR %>%
  filter(outcome == "Forced out") %>% 
  pull(LOR)

LOR_left <- LOR %>%
  filter(outcome == "Left voluntarily") %>%
  pull(LOR)

ROPE_calculation <- function(scale, LOR_data){
  round(sum(LOR_data > -log(scale) & LOR_data < log(scale)) / length(LOR_data), 3)
}

cat(
  paste(
    paste("Probability that |LOR|<2 for % bees leaving voluntarily:", ROPE_calculation(2, LOR_left)), "\n",
    paste("Probability that |LOR|<2 for % bees forced out:", ROPE_calculation(2, LOR_forced)), sep = "")
)
```



### Stats associated with Figure 2C

Here, we derive the result presented in prose in the Results, regarding the effect of LPS on the proportion of bees that left the hive by force. 

```{r}
difference_prop_forced_out_LOR <- plotting_data %>%
  spread(outcome, prop) %>%
  mutate(prop_leavers_that_were_forced_out = 
           `Forced out` / (`Forced out` + `Left voluntarily`)) %>%
  select(posterior_sample, 
         treatment, 
         prop_leavers_that_were_forced_out)  %>%
  spread(treatment, 
         prop_leavers_that_were_forced_out) %>% 
  mutate(difference_prop_forced_out_LOR = get_log_odds(Ringers, LPS)) %>%
  pull(difference_prop_forced_out_LOR)

hypothesis(difference_prop_forced_out_LOR, 
           "x > 0", alpha = 0.05)$hypothesis
```

